Solving non smooth dynamics using ADMM

### 2020-10-30 2021/ADMM_journal/author.tex
Update paper ADMM

\documentclass[1p]{elsarticle}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
%\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{tabularx}
%\usepackage{algorithm}
%\usepackage{algpseudocode}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage{array}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage[title]{appendix}
\usepackage{empheq}

@@

\def\avect#1{{\boldsymbol{#1}}}
\def\amatr#1{{\boldsymbol{#1}}}

\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\mat}[1]{#1}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

@@

\begin{subequations}
	\begin{empheq}[left=\empheqlbrace]{align}
    M \vect{v} - \vect{k} - D_\epsilon \vect{\gamma}_\epsilon &= 0 \\
    D_{\epsilon}^T \vect{v}  + \vect{b}_\epsilon &= \vect{u}_\epsilon\\
    -\Upsilon^{\circ} \ni \vect{u}_\epsilon  \quad \bot &\quad  \vect{\gamma}_\epsilon \in \Upsilon
	\end{empheq}
	\label{eq:ChronoMCCP}
\end{subequations}

where 
\begin{itemize}
		\item unknown $\vect{v}$ is the speed at the end of the time step, 
		\item unknown $\vect{\gamma}_\epsilon$ is the reaction in contacts and bilateral joints, a vector-signed measure with Lebesgue decomposition in atomic parts (impacts) and continuous parts (continuous reactions),
    \item $M$ is a positive definite, block diagonal matrix containing masses and inertia tensors of the bodies
    \item $D_\epsilon$ is a sparse matrix, the transpose jacobian of all constraints
    \item $\vect{k}$ is a vector containing terms proportional to applied forces, among other things 
    \item $\vect{b}$ is a vector containing constraint stabilization terms, among other things
    \item $\Upsilon$ is the Cartesian product of all cones of admissible constraint forces, $\Upsilon = \bigtimes_i \Upsilon_i$
    \item if a frictional contact is added, $\Upsilon_i \subset \mathbb{R}^3$ is a second order Lorentz cone with aperture proportional to friction coefficient
    \item if a bilateral constraint is added, $\Upsilon_i = \mathbb{R}$ and $\Upsilon^\circ_i = \{0\}$
    \item if a unilateral constraint is added, $\Upsilon_i = \mathbb{R}^+$ and $\Upsilon^\circ_i = \mathbb{R}^-$
    \item $\Upsilon^{\circ}$ is the polar cone, opposite of the dual cone, i.e. $\Upsilon^{*} = -\Upsilon^\circ$.
\end{itemize}

@@

Introducing the Schur complement 
\[
N=D_{\epsilon}^T M^{-1} D_{\epsilon}
\]
and the vector
\[
\vect{r}_\epsilon = D_{\epsilon}^T M^{-1} \vect{k} + \vect{b}_\epsilon
\]
such that
\[
\vect{u}_\epsilon = N \vect{\gamma}_\epsilon + \vect{r}_\epsilon \, , 
\]
one can also write the problem as the following CCP:
\begin{align}
    -\Upsilon^{\circ} \ni  N \vect{\gamma}_\epsilon + \vect{r}_\epsilon 
    \quad \bot \quad  
    \vect{\gamma}_\epsilon \in \Upsilon
	\label{eq:ChronoCCP}
\end{align}

This CCP corresponds exactly to a first order optimality condition of a convex quadratic program:
\begin{subequations}
	\begin{empheq}[box=\fbox]{align}
	\text{min} \quad & \frac{1}{2} \vect{\gamma}_\epsilon^T N \vect{\gamma}_\epsilon + \vect{r}^T_\epsilon \vect{\gamma}_\epsilon \\
	\text{s.t.} \quad & \vect{\gamma}_\epsilon \in \Upsilon
	\end{empheq}
	\label{eq:ChronoCCP_min}
\end{subequations}
%
and in fact, the CCP \eqref{eq:ChronoCCP} can be written in the more conventional language of the KKT optimality conditions on \textit{dual variables} $\vect{y}$ multipliers, for $\vect{y}=-\vect{u}_\epsilon$ and \textit{primal variables} $\vect{\gamma}_\epsilon$:
\begin{subequations}
	\begin{align}
    N \vect{\gamma}_\epsilon + \vect{r}_\epsilon + I \vect{y} &= \vect{0} \\
    \vect{\gamma}_\epsilon &= \vect{z} \\
    \Upsilon \ni \vect{z}  \quad \bot &\quad \vect{y} \in \Upsilon^\circ  
	\end{align}
	\label{eq:ChronoCCP_kkt}
\end{subequations}
After the convex program \eqref{eq:ChronoCCP_min} is solved by ADMM, one can compute $\vect{v} = M^{-1}( \vect{k} + \bar{D}_\epsilon \vect{\gamma}_\epsilon)$ with a quick postprocessing step.

@@

Also, in sake of highest generality, in presence of finite elements one might need to use tangent stiffness matrices $K$ and damping matrices $D$ to accommodate an implicit integration scheme for stiff elements, and this means that the Schur complement would be computed as $N=D_{\epsilon}^T H^{-1} D_{\epsilon}$. Here $H$, taking the place of the original $M$ matrix, is rather a linear combination of $M$, $K$, $D$ (depending on the integration scheme) and in general is not diagonal anymore.

% TO ADD IN EXTENDED VERSION
%This means that one cannot easily compute the Schur complement $N$, even in factored form, and this has some consequences on the implementation of the ADMM, that in fact we will rewrite in an unrolled form.

@@

For variables $(\vect{x},\vect{z}) \in \mathbb{R}^m \times \mathbb{R}^n$, the ADMM methods solve constrained optimization problems with separable structure
\begin{subequations}
	\begin{align}
    \text{min} \quad  & f(\vect{x}) + g(\vect{z}) \\
	  \text{s.t.} \quad &  A \vect{x} + B \vect{z} - \vect{b} = \vect{0}
	\end{align}
	\label{eq:admm}
\end{subequations}
by introducing an augmented Lagrangian
\begin{align}
\mathcal{L}_{\rho}(\vect{x},\vect{z},\vect{y})= f(\vect{x}) + g(\vect{z}) + \vect{y}^T (A \vect{x} + B \vect{z} - \vect{b}) + \frac{\rho}{2} \norm{A \vect{x} + B \vect{z} - \vect{b}}
\end{align}
and by iterating over two minimization problems as in the following loop:
\begin{align}
 \vect{x}^{k+1} &\in \text{argmin} \mathcal{L}_{\rho}(\vect{x},\vect{z}^k,\vect{y}^k) \label{eq:lag1}\\
 \vect{z}^{k+1} &\in \text{argmin} \mathcal{L}_{\rho}(\vect{x}^{k+1},\vect{z},\vect{y}^k) \label{eq:lag2}\\
 \vect{y}^{k+1} &= \vect{y}^{k} + \rho ( A \vect{x}^{k+1} + B \vect{z}^{k+1} - \vect{b} )
 \label{eq:admm_iters}
\end{align}
Although the convergence is fast in the first iterations, it tends to deteriorate later as in fixed-point iterations; however in practical scenarios where loose tolerances can be accepted, the method proves to be very efficient and robust.

@@

In order to rewrite \eqref{eq:ChronoCCP_min} as a sum of two functions as in \eqref{eq:admm}, we introduce the auxiliary variables 
$(\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}})$, 
we state the $\vect{\gamma}_\epsilon \in \Upsilon$ constraint by adding a non-smooth penalty function given by the indicator function 
$\mathcal{I}_\Upsilon (\vect{\gamma}_\epsilon)$ 
where $\mathcal{I}_\Upsilon(\vect{\gamma}_\epsilon)=0$ for $\vect{\gamma}_\epsilon \in \Upsilon$ and $=\infty$ otherwise, another indicator function 
$\mathcal{I}_{\vect{\gamma}_\epsilon = \vect{z}}(\vect{\gamma}_\epsilon,\vect{z})$,
and finally we have
\begin{subequations}
	\begin{align}
    \text{min} \quad & \frac{1}{2} \tilde{\vect{\gamma}}_\epsilon^T N \tilde{\vect{\gamma}}_\epsilon + \vect{r}^T_\epsilon \tilde{\vect{\gamma}}_\epsilon  +  \mathcal{I}_{\tilde{\vect{\gamma}}_\epsilon = \tilde{\vect{z}}}(\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}) +  \mathcal{I}_\Upsilon(\vect{\gamma}_\epsilon)  \\
	  \text{s.t.} \quad & (\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}})  =  (\vect{\gamma}_\epsilon,\vect{z})
	\end{align}
	\label{eq:admm_mod}
\end{subequations}

We can write the augmented Lagrangian introducing two step size parameters $\sigma$ and $\rho$ and two vector of multipliers $\vect{w}$ and $\vect{y}$, obtaining:
\begin{align}
\mathcal{L}_{\sigma,\rho} \left((\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}),(\vect{\gamma}_\epsilon,\vect{z}),(\vect{y},\vect{w})\right) &= 
\frac{1}{2} \tilde{\vect{\gamma}}_\epsilon^T N \tilde{\vect{\gamma}}_\epsilon + \vect{r}^T_\epsilon \tilde{\vect{\gamma}}_\epsilon  
+  \mathcal{I}_{\tilde{\vect{\gamma}}_\epsilon = \tilde{\vect{z}}}(\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}) 
+  \mathcal{I}_\Upsilon(\vect{\gamma}_\epsilon) \nonumber \\
+ \vect{w}^T (\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon) 
&+ \frac{\sigma}{2} \norm{\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon} 
+ \vect{y}^T (\tilde{\vect{z}} - \vect{z}) 
+ \frac{\rho}{2} \norm{\tilde{\vect{z}} - \vect{z}}
\end{align}
that is also, using the property $\vect{y}^T\vect{r} + \frac{\rho}{2}\norm{\vect{r}} = \frac{\rho}{2}\norm{\vect{r}+\frac{1}{\rho} \vect{y}} - \frac{1}{2\rho} \norm{\vect{y}}$, the following
\begin{align}
\mathcal{L}_{\sigma,\rho} \left((\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}),(\vect{\gamma}_\epsilon,\vect{z}),(\vect{y},\vect{w})\right) &= 
\frac{1}{2} \tilde{\vect{\gamma}}_\epsilon^T N \tilde{\vect{\gamma}}_\epsilon + \vect{r}^T_\epsilon \tilde{\vect{\gamma}}_\epsilon  
+  \mathcal{I}_{\tilde{\vect{\gamma}}_\epsilon = \tilde{\vect{z}}}(\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}) 
+  \mathcal{I}_\Upsilon(\vect{\gamma}_\epsilon) \nonumber \\
+ \frac{\sigma}{2} \norm{\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon + \frac{1}{\sigma} \vect{w}} 
&- \frac{1}{2\sigma} \norm{\vect{w}}
+ \frac{\rho}{2} \norm{\tilde{\vect{z}} - \vect{z} + \frac{1}{\rho} \vect{y}}
- \frac{1}{2\rho} \norm{\vect{y}}
\end{align}

>>>>

\documentclass[final,3p]{elsarticle}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
%\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{tabularx}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage{array}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage[title]{appendix}
\usepackage{empheq}

@@

%\def\avect#1{{\boldsymbol{#1}}}
\def\amatr#1{{#1}}
\newcommand{\vect}[1]{\bm{#1}}
%\newcommand{\mat}[1]{#1}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays

\newtheorem*{definition*}{Definition}
\newtheorem{definition}{Definition}
\newtheorem*{remark*}{Remark}
\newtheorem{remark}{Remark}
\newtheorem*{theorem*}{Theorem}
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{corollary*}{Corollary}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Definitions}

In this section we list some basic definitions that we will use in the rest of the paper.

\begin{definition}
A \textit{second order cone}, also Lorentz cone, is a self-dual self-scaled symmetric cone defined as
\begin{equation}
\label{eq:lorentzcone}
\mathcal{K} 
= \left \{ (x_0,\vect{x}_1) \in \mathbb{R} \times \mathbb{R}^{p-1}: \norm{\vect{x}_1}_2 \leq x_0  \right \} 
\; =   -\mathcal{K}^*.
\end{equation}
\end{definition}

\begin{definition}
The \textit{dual cone} $\mathcal{K}^*$ of a set $\mathcal{K}$ in a vector space equipped with an inner product, is always convex and is defined as
\begin{equation}
\label{eq:dualcone}
\mathcal{K}^* 
= \left \{ \vect{y} \in \mathbb{R}^n: \left\langle \vect{y}, \vect{x} \right\rangle \geq 0 \quad \forall \vect{x} \in \mathcal{K}  \right \}.
\end{equation}
\end{definition}

\begin{definition}
The \textit{polar cone} is defined as
\begin{equation}
\label{eq:polarcone}
\mathcal{K}^\circ 
= \left \{ \vect{y} \in \mathbb{R}^n: \left\langle \vect{y}, \vect{x} \right\rangle \leq 0 \quad \forall \vect{x} \in \mathcal{K}  \right \} 
\; =   -\mathcal{K}^*.
\end{equation}
\end{definition}

\begin{definition}
The \textit{normal cone} to a closed convex set $\mathcal{K}$ at the point $\vect{x}\in\mathcal{K}$ is
closed and convex and is defined as
\begin{align}
\label{eq:normalcone}
\mathcal{N}_{\mathcal{K}}(\vect{x}) 
&= \{ \vect{y} \in \mathbb{R}^n : \left\langle \vect{y}, \vect{x} - \vect{z} \right\rangle \geq 0, \forall \vect{z} \in \mathcal{K} \} 
\end{align}
\end{definition}

\begin{definition}
The \textit{indicator function} of a subset $\mathcal{A} \in \mathcal{E}$ is a scalar function $I: \mathcal{E} \mapsto \mathbb{R}$
defined as:
\begin{equation}
\label{eq:indicator}
{I}_{\mathcal{A}}(\vect{x}) = 
\left\{ 
\begin{array}[pos]{l}
0     \;  \text{if} \; \vect{x} \in \mathcal{A} \\
\infty   \; \text{if} \; \vect{x} \notin \mathcal{A}
\end{array}
\right.
\end{equation}
\end{definition}

\begin{remark*}
The indicator function of any closed set is lower semicontinuous.
\end{remark*}

\begin{definition}
The \textit{subdifferential} $\partial f(\vect{x}_0)$ at $\vect{x}_0$ of a convex, possibly non differentiable scalar function $f: \mathbb{R}^n \mapsto \mathbb{R}$, is the closed convex set of all subgradients $\vect{g}$ at $\vect{x}_0$:
\begin{equation}
\label{eq:subdifferential}
\partial f(\vect{x}_0) = \left\{ \vect{g} : 
  f(\vect{x}) -  f(\vect{x}_0) \geq \left\langle \vect{g}, (\vect{x} - \vect{x}_0) \right\rangle \; \forall \vect{x} \in \mathcal{E} \right\}.
\end{equation}
\end{definition}

\begin{remark*}
The indicator function of any closed set is lower semicontinuous.
\end{remark*}

\begin{remark*}
The subdifferential is a set valued function $\partial f(\vect{x}): \mathbb{R}^n \rightrightarrows \mathbb{R}^n$ in general, but as a special case, if $f(\vect{x})$ is differentiable, $\partial f(\vect{x}) = \left\{ \nabla f(\vect{x}) \right\}$.
\end{remark*}

\begin{remark*}
The subdifferential of an indicator function of a convex set is also its normal cone:
%
\begin{equation}
\label{eq:sub_ind}
\partial {I}_{\mathcal{K}}(\vect{x}) = \mathcal{N}_{\mathcal{K}}(\vect{x}).
\end{equation}
\end{remark*}

\begin{definition}
The \textit{projection} of $\vect{y}$ on a nonempty closed convex set $\mathcal{E}$ is a $\mathbb{R}^n \rightarrow \mathbb{R}^n$ mapping defined as:
%
\begin{align}
\label{eq:projector}
\Pi_{\mathcal{E}}(\vect{y}) = \mathrm{argmin}_{\vect{x} \in \mathcal{E}} \norm{ \vect{x} - \vect{y} }^2_2
\end{align}
%
If $f(\vect{y})$ is the indicator function $I_{\mathcal{E}}(\vect{y})$ to a nonempty closed convex set ${\mathcal{E}}$, defined as in \eqref{eq:indicator}, then $\mathrm{prox}_{f}(\vect{y})$ is the orthogonal projection onto that set:
%
\begin{align}
\label{eq:projectorindicator}
\mathrm{prox}_{I_{\mathcal{E}}}(\vect{y})  = \Pi_{\mathcal{E}}(\vect{y}) 
\end{align}
\label{def:projection}
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{subequations}
	\begin{empheq}[left=\empheqlbrace]{align}
    H \vect{v} - \vect{k} - D_\epsilon \vect{\gamma}_\epsilon &= 0   \label{eq:ChronoMCCP_a} \\
    D_{\epsilon}^T \vect{v}  + \vect{b}_\epsilon &= \vect{u}_\epsilon   \label{eq:ChronoMCCP_b} \\
    -\Upsilon^{\circ} \ni \vect{u}_\epsilon  \quad \bot &\quad  \vect{\gamma}_\epsilon \in \Upsilon  \label{eq:ChronoMCCP_c}
	\end{empheq}
	\label{eq:ChronoMCCP}
\end{subequations}

where 
\begin{itemize}
		\item unknown $\vect{v}$ is the speed at the end of the time step $h$
		\item unknown $\vect{\gamma}_\epsilon$ is the reaction in contacts and bilateral joints, a vector-signed measure with Lebesgue decomposition in atomic parts (impacts) and continuous parts (continuous reactions)
    \item $H$ is a positive definite matrix containing $M$, the block-diagonal matrix of the masses and inertia tensors of the bodies; if stiff
		loads are added too (ex. when finite elements come into play) it becomes a block-sparse matrix including the tangent stiffness and damping matrices,
		for instance $H=M-{h^2}\nabla_q\vect{f}-{h}\nabla\vect{f}$ in case of the time stepper in \cite{TasoraAnitescuCMAME10}
    \item $D_\epsilon$ is a sparse matrix, the transpose jacobian of all constraints
    \item $\vect{k}$ is a vector containing terms proportional to applied forces $\vect{f}$
    \item $\vect{b}$ is a vector containing constraint stabilization terms, among other things
    \item $\Upsilon$ is the Cartesian product of all cones of admissible constraint forces, $\Upsilon = \bigtimes_i \Upsilon_i$
    \item if a frictional contact is added, $\Upsilon_i \subset \mathbb{R}^3$ is a second order Lorentz cone with aperture proportional to friction coefficient
    \item if a bilateral constraint is added, $\Upsilon_i = \mathbb{R}$ and $\Upsilon^\circ_i = \{0\}$
    \item if a unilateral constraint is added, $\Upsilon_i = \mathbb{R}^+$ and $\Upsilon^\circ_i = \mathbb{R}^-$
    \item $\Upsilon^{\circ}$ is the polar cone, opposite of the dual cone, i.e. $\Upsilon^{*} = -\Upsilon^\circ$.
\end{itemize}

@@

Introducing the Schur complement 
\begin{align}
N=D_{\epsilon}^T H^{-1} D_{\epsilon}
\label{eq:schur_n}
\end{align}
and the vector
\begin{align}
\vect{r}_\epsilon = D_{\epsilon}^T H^{-1} \vect{k} + \vect{b}_\epsilon
\label{eq:schur_r}
\end{align}
such that
\[
\vect{u}_\epsilon = N \vect{\gamma}_\epsilon + \vect{r}_\epsilon \, , 
\]
one can also write the problem as the following CCP:
\begin{align}
    -\Upsilon^{\circ} \ni  N \vect{\gamma}_\epsilon + \vect{r}_\epsilon 
    \quad \bot \quad  
    \vect{\gamma}_\epsilon \in \Upsilon
	\label{eq:ChronoCCP}
\end{align}

This CCP corresponds exactly to a first order optimality condition of a convex quadratic program:
\begin{subequations}
	\begin{empheq}[box=\fbox]{align}
	\text{min} \quad & \frac{1}{2} \vect{\gamma}_\epsilon^T N \vect{\gamma}_\epsilon + \vect{r}^T_\epsilon \vect{\gamma}_\epsilon \\
	\text{s.t.} \quad & \vect{\gamma}_\epsilon \in \Upsilon
	\end{empheq}
	\label{eq:ChronoCCP_min}
\end{subequations}
%
and in fact, the CCP \eqref{eq:ChronoCCP} can be written in the more conventional language of the KKT optimality conditions on \textit{dual variables} $\vect{y}$ multipliers, for $\vect{y}=-\vect{u}_\epsilon$, and \textit{primal variables} $\vect{\gamma}_\epsilon$:
\begin{subequations}
	\begin{align}
    N \vect{\gamma}_\epsilon + \vect{r}_\epsilon + I \vect{y} &= \vect{0} \\
    \vect{\gamma}_\epsilon &= \vect{z} \\
    \Upsilon \ni \vect{z}  \quad \bot &\quad \vect{y} \in \Upsilon^\circ  
	\end{align}
	\label{eq:ChronoCCP_kkt}
\end{subequations}
After the convex program \eqref{eq:ChronoCCP_min} is solved by ADMM, one can compute 
\begin{align}
\vect{v} = H^{-1}( \vect{k} + \bar{D}_\epsilon \vect{\gamma}_\epsilon)
\label{eq:v_post}
\end{align}
with an immediate step.

@@

Also, in sake of highest generality, in presence of finite elements one might need to use tangent stiffness matrices $K$ and damping matrices $D$ to accommodate an implicit integration scheme for stiff elements, and this means that the Schur complement would be computed as $N=D_{\epsilon}^T H^{-1} D_{\epsilon}$. Here $H$ is a linear combination of $M$, $K$, $D$ (depending on the integration scheme) and in general is not diagonal anymore.

@@

For variables $(\vect{x},\vect{z}) \in \mathbb{R}^m \times \mathbb{R}^n$, the ADMM methods solve constrained optimization problems with separable structure
\begin{subequations}
	\begin{align}
    \text{min} \quad  & f(\vect{x}) + g(\vect{z}) \\
	  \text{s.t.} \quad &  A \vect{x} + B \vect{z} - \vect{b} = \vect{0}
	\end{align}
	\label{eq:admm}
\end{subequations}
where $f : \mathbb{R}^{n_x}\rightarrow\overline{\mathbb{R}}$ and  $g : \mathbb{R}^{n_z}\rightarrow\overline{\mathbb{R}}$ are proper convex lower-semicontinuous functions,
and $A \in \mathbb{R}^{n_x \times n_y}$, $B \in \mathbb{R}^{n_z \times n_y}$ are linear operators.

The optimality conditions require that primal feasibility
\begin{align}
A \vect{x}^\star + B \vect{z}^\star - \vect{b} = \vect{0} \label{eq:primalfeasibility1}
\end{align}
and dual feasibility
\begin{subequations}
	\begin{align}
     0 \in \partial	f(\vect{x}^\star) + A^T \vect{y}^\star \label{eq:dualfeasibility1}\\
	   0 \in \partial	g(\vect{z}^\star) + B^T \vect{y}^\star \label{eq:dualfeasibility2}
	\end{align}
	\label{eq:admm}
\end{subequations}
must hold at the solution $(\vect{x}^\star,\vect{z}^\star,\vect{y}^\star)$, saddle point of the Lagrangian with dual variables $\vect{y} \in \mathbb{R}^{n_y}$.

By introducing an augmented Lagrangian
\begin{align}
\mathcal{L}_{\rho}(\vect{x},\vect{z},\vect{y})= f(\vect{x}) + g(\vect{z}) + \vect{y}^T (A \vect{x} + B \vect{z} - \vect{b}) + \frac{\rho}{2} \norm{A \vect{x} + B \vect{z} - \vect{b}}_2^2
\end{align}
the ADMM method converges to the solution $(\vect{x}^\star,\vect{z}^\star,\vect{y}^\star)$ by iterating over two distinct minimization problems as in the following loop:
\begin{align}
 \vect{x}^{k+1} &\in \argmin_{\vect{x}} \mathcal{L}_{\rho}(\vect{x},\vect{z}^k,\vect{y}^k) \label{eq:lag1}\\
 \vect{z}^{k+1} &\in \argmin_{\vect{z}} \mathcal{L}_{\rho}(\vect{x}^{k+1},\vect{z},\vect{y}^k) \label{eq:lag2}\\
 \vect{y}^{k+1} &= \vect{y}^{k} + \rho ( A \vect{x}^{k+1} + B \vect{z}^{k+1} - \vect{b} )
 \label{eq:admm_iters}
\end{align}
The convergence of such method has rate $\mathcal{O}\left(\frac{1}{k}\right)$.
Although the convergence is fast in the first iterations, it tends to deteriorate later as in most fixed-point iterations; however in practical scenarios where loose tolerances can be accepted, even this basic formulation of ADMM proves to be efficient and robust.

@@

In order to rewrite \eqref{eq:ChronoCCP_min} as a sum of two functions as in \eqref{eq:admm}, we introduce 
$\vect{\gamma}_\epsilon \in \mathbb{R}^{n_c}$ as primal variables, 
$\vect{z} \in \mathbb{R}^{n_c}$ as auxiliary variables, 
we add $\vect{\gamma}_\epsilon - \vect{z} = \vect{0}$ as the linear constraint, 
we reformulate the $\vect{\gamma}_\epsilon \in \Upsilon$ constraint by adding a non-smooth penalty function given by the indicator function 
$\mathcal{I}_\Upsilon (\vect{z})$,
and finally we have
\begin{subequations}
	\begin{align}
    \text{min} \quad & \frac{1}{2} \vect{\gamma}_\epsilon^T N \vect{\gamma}_\epsilon + \vect{r}^T_\epsilon \vect{\gamma}_\epsilon  
		+  \mathcal{I}_\Upsilon(\vect{z})  \\
	  \text{s.t.} \quad & \vect{\gamma}_\epsilon =  \vect{z}
	\end{align}
	\label{eq:admm_mod}
\end{subequations}

We can write the augmented Lagrangian introducing a step size parameter $\rho$ and a vector of dual variables $\vect{y}$, obtaining:
\begin{align}
\mathcal{L}_{\rho} \left(\vect{\gamma}_\epsilon,\vect{z},\vect{y} \right) &= 
\frac{1}{2} \vect{\gamma}_\epsilon^T N \vect{\gamma}_\epsilon + \vect{r}^T_\epsilon \vect{\gamma}_\epsilon  
+  \mathcal{I}_\Upsilon(\vect{z}) \nonumber \\
&+ \vect{y}^T (\vect{\gamma}_\epsilon - \vect{z}) 
+ \frac{\rho}{2} \norm{\vect{\gamma}_\epsilon - \vect{z}}_2^2
\end{align}
that is also, using the property $\vect{y}^T\vect{r} + \frac{\rho}{2}\norm{\vect{r}}_2^2 = \frac{\rho}{2}\norm{\vect{r}+\frac{1}{\rho} \vect{y}}_2^2 - \frac{1}{2\rho} \norm{\vect{y}}_2^2$, the following
\begin{align}
\mathcal{L}_{\rho} \left(\vect{\gamma}_\epsilon,\vect{z},\vect{y} \right) &= 
\frac{1}{2} \vect{\gamma}_\epsilon^T N \vect{\gamma}_\epsilon + \vect{r}^T_\epsilon \vect{\gamma}_\epsilon  
+  \mathcal{I}_\Upsilon(\vect{z}) \nonumber \\
&+ \frac{\rho}{2} \norm{\vect{\gamma}_\epsilon - \vect{z} + \frac{1}{\rho} \vect{y}}_2^2
- \frac{1}{2\rho} \norm{\vect{y}}_2^2
\label{eq:augmentedlagrangian}
\end{align}

The first step of ADMM requires to compute 
\begin{align}
\vect{\gamma}_\epsilon^{k+1} &= \argmin_{\vect{\gamma}_\epsilon} \mathcal{L}_{\rho} \left( \vect{\gamma}_\epsilon,\vect{z}^k,\vect{y}^k \right) \\
 &= \argmin_{\vect{\gamma}_\epsilon} \frac{1}{2} \vect{\gamma}_\epsilon^T N \vect{\gamma}_\epsilon + \vect{r}^T_\epsilon \vect{\gamma}_\epsilon  
+ \frac{\rho}{2} (\vect{\gamma}_\epsilon - \vect{z}^k + \frac{1}{\rho} \vect{y}^k)^T(\vect{\gamma}_\epsilon - \vect{z}^k + \frac{1}{\rho} \vect{y}^k)
- \frac{1}{2\rho} \norm{\vect{y}^k}_2^2 \\
 &= \argmin_{\vect{\gamma}_\epsilon} \frac{1}{2} \vect{\gamma}_\epsilon^T (N+ \rho I) \vect{\gamma}_\epsilon 
   + (\vect{r}_\epsilon - \rho \vect{z}^k + \vect{y}^k)^T \vect{\gamma}_\epsilon + C^k 
\end{align}

This is an unconstrained convex quadratic program whose KKT optimality conditions lead to the following linear problem:
\begin{align}
    \left[ N + \rho I \right] \vect{\gamma}_\epsilon^{k+1} =  \rho \vect{z}^k - \vect{r} - \vect{y}^k
		\label{eq:admm_1b}
\end{align}

%Optionally, following the idea of \cite{Stellato2020}, one could also introduce a small regularization parameter $\sigma$, that helps in case $N$ is near singular and $\rho$ is very small:
%\begin{align}
    %\left[ N + (\sigma+\rho) I \right] \tilde{\vect{\gamma}}_\epsilon = \sigma \vect{\gamma}_\epsilon^k + \rho \vect{z}^k - \vect{r} - \vect{y}^k
		%\label{eq:admm_1breg}
%\end{align}

The linear system \eqref{eq:admm_1b} can be solved as it is, but in our case it would be better to exploit the fact that the Schur matrix $N$ is a product $D_\epsilon^T H^{-1} D_\epsilon$. We would like to avoid computing $H^{-1}$, even storing a precomputed inverse for all iterations would be unpractical because often dense (except when $H$ is a diagonal mass matrix $M$). So we propose to replace \eqref{eq:admm_1b} with an equivalent saddle point problem:

\begin{theorem}
\label{th:schurtokkt}
Let $\vect{\gamma}_\epsilon^{k+1}$ be a solution to \eqref{eq:admm_1b}, then it is also a solution to
%
\begin{subequations}
	\begin{align}
    \begin{bmatrix}
		 H   & D_\epsilon \\
		 D_\epsilon^T & - \rho I
		\end{bmatrix}
		\begin{Bmatrix}
		 \vect{v}^{k+1}   \\
		 -\vect{\gamma}_\epsilon^{k+1} 
		\end{Bmatrix}
		=
		\begin{Bmatrix}
		 \vect{k} \\
		 -\vect{b}_\epsilon + \rho \vect{z}^k -\vect{y}^k 
		\end{Bmatrix}
	\end{align}
	\label{eq:admm_1c}
\end{subequations}
%
\end{theorem}

\begin{proof}
Performing the multiplication of the upper part of the saddle point matrix one has 
$H \vect{v}^{k+1} - D_\epsilon \vect{\gamma}_\epsilon^{k+1} = \vect{k}$,
pre-multiplying by $H^{-1}$ one gets 
$\vect{v}^{k+1} = H^{-1} D_\epsilon \vect{\gamma}_\epsilon^{k+1} + H^{-1} \vect{k}$. 
The multiplication of the lower part gives 
$D_\epsilon^T \vect{v}^{k+1} + \rho I \vect{\gamma}_\epsilon^{k+1} = -\vect{b}_\epsilon + \rho \vect{z}^k -\vect{y}^k $, 
hence after substitution of $\vect{v}^{k+1}$ one gets
$D_\epsilon^T H^{-1} D_\epsilon \vect{\gamma}_\epsilon^{k+1}  + D_\epsilon^T H^{-1} \vect{k} + \rho I \vect{\gamma}_\epsilon^{k+1} = -\vect{b}_\epsilon + \rho \vect{z}^k -\vect{y}^k$. 
Recalling the definition of $N$ in \eqref{eq:schur_n} and the definition of $\vect{r}_\epsilon$ in \eqref{eq:schur_r}, this can be written also as 
$(N + \rho I) \vect{\gamma}_\epsilon^{k+1} =  \rho \vect{z}^k - \vect{r} - \vect{y}^k$. 
\end{proof}

\begin{corollary}
The linear system \eqref{eq:admm_1c} introduces an auxiliary variable $\vect{v} \in \mathbb{R}^{n_v}$, however the matrix is very sparse and does not require storing or computing any $H^{-1}$ matrix as we should do if using \eqref{eq:admm_1b}.
\end{corollary}

\begin{corollary}
A positive side effect of this approach is that its auxiliary variable $\vect{v}$ is also the velocity term in the original mixed CCP 
\eqref{eq:ChronoMCCP}, so one does not need to compute it as $\vect{v}=H^{-1}(\vect{k} + D_\epsilon \vect{\gamma}_\epsilon )$ after the iteration converged because it is already a byproduct of the linear solver in \eqref{eq:admm_1c}.
\end{corollary}
